Theory about Normal Basis on $F(2^k)$

### 2016-12-09 dissertation/ONB.tex
dissertation v2:abstract & appendix update 2

% LaTeX file for a 1 page document
\documentclass[12pt]{article}
%\usepackage{e-jc}

@@

\date{Last Update: Apr 17, 2013}

@@

\textbf{Preliminary to Lenstra's Algorithm}
\begin{itemize}
\item[-] \textbf{Lemma 4}\ \ \ \ For an arbitrary element $\theta \in F_{q^n}$ that $Ord_\theta(x) \neq x^n - 1$,
let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.\
\item[-] \textbf{Lemma 5}\ \ \ \ Same $\theta$ and $g(x)$ defined as last lemma. Assume there exists a solution 
$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Then there exists a non-zero element $\eta$ such that\\
$g(\sigma)\eta = 0$, and\\
$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$.

@@

Suppose we take only one polynomial only contains T(next state), like\\
\hspace{8mm}\par
$ideal\ G = T^2 + (1+X) \cdot T + X$\\
\hspace{8mm}\par
Since we get G from GB based image function, so G is already a Groebner Basis itself.
Considering Theorem 1 $\sim$ 3, it's easy to do ideal operation if its generator is only one polynomial.
So using $G+G'$ we can get intersection, using $G \cdot G'$ we can get union. If we take V = Universe = $<vanishing\ polynomial>$, we can get ideal quotient $I(V):I(W)$ as complementary set for
specific varieties.

\hspace{8mm}\par
\textbf{More about Ideal Quotient}
\begin{itemize}
\item[-] \textbf{Definition 1}\ \ \ \ If I,J are ideals in $k[x_1,...,x_n]$, then $I:J$ is the set \\ \indent $\{f\in k[x_1,...,x_n]:fg\in I\ for \ all\ g\in J\}$
\end{itemize} \par
Why we can get complementary set $\overline {reached}$ through this?\par
First, we only care about varieties. Say we redefine the "equal" as $V(I)==V(J)\Leftrightarrow I == J$.\par
Second, we only care about the ideal/GB {\bf G} contains only 1 generator f. Say this polynomial f is a function about next state (word level) T, then $f(T=V(\bf{G}))==0$. To ensure then unique of generator f, we'll reduce f (or say {\bf G}) every time.\par
Third, for ideal quotient ${\bf U:G}$ here, we need to find f from $g~f==vanish$. This means $<f>\bigcup <g>==<vanish>$. How can we develop a division algorithm to get exactly a factor without any remainders? Let's see 3 examples following:\\
\hspace{8mm}\par
\indent $T^2+(1+X)\cdot T+X$\\
\hspace{8mm}\par
\indent $T^2+T$\\
\hspace{8mm}\par
\indent $T^2+1$\\

>>>>

% LaTeX file for a 1 page document
\documentclass[12pt]{article}
%\usepackage{e-jc}
\usepackage{amsfonts}

@@

\date{Last Update: May 14, 2013}

@@

\textbf{Preliminary to Lenstra's Algorithm}
\begin{itemize}
\item[-] \textbf{Lemma 4}\ \ \ \ For an arbitrary element $\theta \in F_{q^n}$ that $Ord_\theta(x) \neq x^n - 1$,
let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.\
\item[-] \textbf{Lemma 5}\ \ \ \ Same $\theta$ and $g(x)$ defined as last lemma. Assume there exists a solution 
$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Then there exists a non-zero element $\eta$ such that\\
$g(\sigma)\eta = 0$, and\\
$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$.\\

@@

\textbf{Lenstra's Algorithm}\\
Step 1: Take an arbitrary element $\theta \in F_{q^n}$, determine $Ord_\theta(x)$.\\
Step 2: If $Ord_\theta(x) = x^n - 1$ then algorithm ends.\\
Step 3: Calculate $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.\\
Step 4: Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to step 2;
	otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
	replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, then go to step 2.\\

\section{Optimal Normal Basis Characterization}
\textbf{Definition of Optimal Normal Basis}\\
To define optimal normal basis, it is necessary to find out the minimum number
of terms in multiplication table(defined below).
\textit{Theorem 3}\ \ If N is a normal basis for $F_{p^n}$ with multiplication table $M_T$, then non-zero terms in 
matrix $C_N\leq 2n-1$.\\
\textit{Proof}\ \ Let $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}$, denote $\beta^{p^i}$ by $\beta_i$. 
Then $\sum_{i=0}^{n-1} \beta_i = trace \beta$. \\
Denote $trace \beta$ with $b$, consider $n\times n$ matrix (multiplication table for lowest digit) $M_0$. Then\\
$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$.\\
Therefore, the sum of all rows in $M_0$ is an n-tuple with $b$ as the first element and zeros elsewhere.
So the first column must contain at least 1 nonzero element, and other columns must get a sum of zero 
while keeping every row linearly independent so there should be at least 2 non-zero elements each column.\\
For $F_{2^n}$, easy to figure out the first row must contain a "1" because $\beta \beta^{2^0} = \beta^{2^1}$.
for the rest rows $\beta \beta^{2^i} \neq \beta^{2^{i+1}}$,
so there must be at least 2 "1"s to fulfil the equation.\\
\textbf{Multiplication table}\\
$\beta \left( \begin{array} {lcr}
 \beta \\ 
\beta^2 \\ 
\cdot \\ 
\cdot \\ 
\cdot \\ 
\beta^{2^{n-1}} 
\end{array} \right) = M_T \left( \begin{array} {lcr}
 \beta \\ 
\beta^2 \\ 
\cdot \\ 
\cdot \\ 
\cdot \\ 
\beta^{2^{n-1}} 
\end{array} \right)$. \\

Complexity $C_N \geq 2n - 1$.\\

\textbf{ONB Existance Theory}\\
Key words: k-th primitive root of unity, Euler's criterion, quadratic residues\\
\hspace{8mm}\par
Type I ONB: n+1 is prime and q is primitive in $\mathbb{Z}_{n+1}$, then the n nonunit (n+1)th roots of unity form ONB.\\
\hspace{8mm}\par
Type II ONB: 2n+1 is prime, if (1) 2 is primitive in $\mathbb{Z}_{2n+1}$, OR \\
(2) $2n + 1 \equiv 3(mod 4)$ and 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$.\\
Then $\alpha = \gamma + \gamma^{-1}$ generates ONB, where $\gamma$ is a primitive (2n+1)th root if unity.\\

\textbf{Low-complexity Normal Basis design}\\
N/A.\\

\section{Other NB Properties \& Problems}
\textbf{N-poly}\\
N-poly is irreducible polynomial with linearly independent roots. Normal basis is a set of roots of N-poly.
\begin{itemize}
\item[-] \textbf{Corollary 2}\ \ \ \ For irreducible polynomial $f(x) = x^n + a_1x^{n-1} + ... + a_n \in F_{2^n}[x]$, 
		that is an N-poly if and only if $a_1 \neq 0$.
\end{itemize}

\textbf{Problems and Discussion}\\
\hspace{8mm}\par
Prob1: how to prove $\lambda$-Matrix is multiplication table?\\
\indent I think this is a good way to understand the essence of normal basis 
theory. This problem can be rewrite like this:\\
Why there exist a rotating symmetry in $\lambda$-Matrix that
 $\lambda_{0j}^{(k)} = \lambda_{jk}^{(0)}$? Andrew guess it's a Frobenius
 element matrix rotating symmetry involves with field automorphism. Need 
further consultation on this part. How to get this property? Check out the
 last part of my note.\\
\hspace{8mm}\par
Prob2: Proof of Lemma 5.\\
(Already proved. Please check out the note.)\\
\hspace{8mm}\par
Prob3: Fast algorithm for determining $\sigma$-Order.\\
(Do not need FAST algorithms. Just do a $(O(n^2))$ scan.)\\

### 2016-12-09 dissertation/abstract.tex
dissertation v2:abstract & appendix update 2

\begin{center}{\bf ABSTRACT}\end{center}

Verification of sequential circuits remains to be a topic of all time. With increasing size of new
integrated circuits, sequential circuit designers may face much more complicated problems on 
logic errors, unexpected delays and failures on critical path. 
This dissertation addresses the problem of sequential circuit verification at the word-level
and is based on concepts derived from algebraic geometry.

Analyzing the sequential circuits at word level is an efficient way
of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
state-space. We manage to model the verification properties and the gate-level sequential circuit 
implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
techniques from algebraic geometry can be used to reason about the state-space on high-level.
In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
polynomial ideals. Without actually solving the system, the states in state space can be investigated 
by manipulating these ideals.

We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
abstraction-refinement techniques for verification, we propose to investigate a word-level 
analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
not only derive new algorithms and techniques, but will also consider efficient CAD 
implementations to target sequential equivalence and model checking problems. 
% \end{abstract}

>>>>

\begin{center}{\bf ABSTRACT}\end{center}
Sequential equivalence and model checking techniques are widely
employed in hardware and software verification. With the increasing
size and complexity of hardware/software systems, abstraction has
become a key component to address scalability.
This dissertation investigates new abstraction techniques at word-level and their applications 
on sequential circuits verification. By bundling together $k$ bit-level state-variables into
one word-level constraint expression, the state-space can be construed
as a polynomial function over the finite (Galois) field of $2^k$
elements. As long as the state-space are modeled as polynomial ideals
over word-level variables, algebraic geometry techniques including ideal manipulations,
Gr\"obner bases reasoning and finite field algebra are researched to 
facilitate new word-level abstractions.

Algebraic geometry offers a very powerful set of tools to reason about
the properties of the solutions to a system of polynomial
constraints. Moreover, the algebraic model inherently provides a
framework for abstraction. 
The algebraic geometry algorithms (Groebner basis computations) exhibit
high complexity which prevent it from being widely used in conventional techniques. 
Interestingly, recent experience has shown that
by analyzing the topology of circuits in conjunction with Gr\"obner
basis computations helps to identify the structure and symmetry
inherent in the problem, this lower the computational complexity.

Formal verification of sequential circuits requires the analysis of
the underlying finite state machine. Reachability analysis becomes a
fundamental tool for this purpose. We devised traditional traversal algorithms 
using the new abstraction method to perform
the reachability analysis at word-level. 
The abstractions are also exploited on functional
verification of sequential Galois field arithmetic circuits, where conventional 
techniques fail to provide scalability. At last, we propose a new method to extract 
unsatisfiable cores and utilize the information about the cores to 
perform model checking with abstraction refinement.

### 2016-12-09 dissertation/appendix.tex
dissertation v2:abstract & appendix update 2

For a given Frobenius map $\sigma$, corresponding minimal polynomial is restricted:
\begin{Lemma}
The minimal (and characteristic) polynomial for $\sigma$ is $x^n-1$.
\end{Lemma}
\begin{Proof}
Consider Fermat's little theorem in $\F_{p^n}$. Let $\beta$ be an element in the field, then
$$\sigma^n\beta = \beta^{p^n} = \beta \implies \sigma^n-I = 0$$
where $I$ is the identity map. For characteristic polynomial, assume $\exists f(x) = \sum_i f_ix^i \in \F_p[x]$, 
such that the degree of characteristic polynomial is lower than $n$:
$$\sum_if_i\sigma^i = 0~~~\text{and}~~~deg(f(x))<n$$

@@

From lemmas above we can deduce the following corollary:
\begin{Corollary}
An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
\end{Corollary}
\subsection{Construction of General Normal Basis}
\subsection{Bases Conversion}
% General \lambda matrix computation

\section{Optimal Normal Basis}
\label{append:ONB}
\subsection{Construction of Optimal Normal Basis}
\subsection{Optimal Normal Basis Multiplier Design}
% Customized Galois Field IEEE1363-2000
% Multi table <=> \lambda matrix

>>>>

For a given Frobenius map $\sigma$, corresponding minimal polynomial is restricted:
\begin{Lemma}
\label{lem:3}
The minimal (and characteristic) polynomial for $\sigma$ is $x^n-1$.
\end{Lemma}
\begin{Proof}
Consider Fermat's little theorem in $\F_{p^n}$. Let $\beta$ be an element in the field, then
$$\sigma^n\beta = \beta^{p^n} = \beta \implies \sigma^n-I = 0$$
where $I$ is the identity map. For characteristic polynomial, assume $\exists f(x) = \sum_i f_ix^i \in \F_p[x]$, 
such that the degree of characteristic polynomial is lower than $n$:
$$\sum_if_i\sigma^i = 0~~~\text{and}~~~deg(f(x))<n$$

@@

From lemmas above we can deduce the following corollary:
\begin{Corollary}
\label{corol:1}
An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
\end{Corollary}
\begin{Proof}
Normal bases require $\{\beta,\beta^2,\dots,\beta^{2^{n-1}}\}$ to be linearly independent,
which is equivalent to the following expression:
$$\forall f(x)\in \F_p[x],~deg(f(x))<n$$
Furthermore, because there is no annihilators in the $T$-Order, this implies that
$$Ord_{\alpha,\sigma}(x) = x^n-1$$
\end{Proof}

From corollary above we can deduce another form of criterion of normal element:
\begin{Theorem}
\label{thm:1}
Given finite field $\F_{2^n}$, the field characteristic $p = 2$.
Define $t = p^e$ such that $n = kp^e, gcd(k,p) = 1$, so $t=1$ if $n$ is an odd integer. Then 
$x^n - 1$ can be factorized as 
$$(\varphi_1(x)\varphi_2(x)\cdot\cdot\cdot\varphi_r(x))^t$$
Additionally, define $\Phi_i(x) = (x^n - 1)/\varphi_i(x)$. We assert that

An element $\alpha \in 	F_{p^n}$ is a normal element if and only if $\Phi_i(\sigma)\alpha \neq 0, i = 1,2,\dots,r$.
\end{Theorem}
\begin{Proof}
Let us analyze the auxiliary polynomial $\Phi_i(x) = \frac{x^n - 1}{\varphi_i(x)}$ first.
$$\Phi_i(\sigma)\alpha \neq 0 \Leftrightarrow \text{No factor in }x^n-1\text{ annihilates }\alpha$$
On the other hand, according to Lemma \ref{lem:3}, it is a known 
fact that \emph{any annihilator always divide} $x^n-1$. As a result, the only possible situation is
$$Ord_{\alpha,\sigma}(x) = x^n-1$$
Using Corollary \ref{corol:1} we deduce that $\alpha$ is a normal element.
\end{Proof}

The normal element identification can also be described from the nullspace perspective:
\begin{Theorem}
\label{thm:2}
Let $W_i$ be the nullspace of $\varphi_{i}^{t} (x)$ and $\widetilde{W_i}$ the
nullspace of $\varphi_{i}^{t-1} (x)$. Let $\overline{W_i}$ be any subspace of
$W_i$ such that $W_i = \overline{W_i}\oplus \widetilde{W_i}$. Then
$$\F_{p^n} = \displaystyle\sum_{i=1}^{r} \overline{W_i}\oplus \widetilde{W_i}$$
is a direct sum where $dim(\overline{W_i}) = d_i$ and $dim(\widetilde{W_i}) = (t-1)d_i$.

Using above setup we define an element $\alpha \in \F_{p^n}$ with 
$\alpha = \sum_{i=1}^{r} (\overline{\alpha_i} + \widetilde{\alpha_i}), \overline{\alpha_i} \in \overline{W_i}, \widetilde{\alpha_i} \in \widetilde{W_i}$,
as a normal element if and only if $\overline{\alpha_i} \neq 0,~ \forall i = 1,2,\dots,r$.
\end{Theorem}

Ultimately, since there always exist at least one element fulfilling requirements in Theorem \ref{thm:1}
or Theorem \ref{thm:2}, we obtain following theorem:
\begin{Theorem}[Normal Basis Theorem over Finite Fields]
There always exists a normal basis of $\F_{p^n}$ over $\F_p$.
\end{Theorem}
\subsection{Construction of General Normal Bases}
After proving the existence of normal bases, the upcoming question is that how to find such a normal basis/element.
For general normal basis identification, there are two methods widely used: the L\"uneburg's algorithm
and Lenstra's algorithm.\\
\textbf{L\"uneburg's Algorithm:}
\begin{enumerate}[{1)}]
\item Randomly pick an element $\alpha$ from $\F_{p^n}$. For each $i = 0,1,\dots,n-1$, 
compute $\sigma$-Order $f_i = Ord_{\alpha^i}(x)$. Then $x^n - 1 = lcm(f_0,f_1,...,f_{n-1})$.
\item Apply factor refinement to set $\{f_0,\dots,f_{n-1}\}$ and obtain $f_i = \prod_{1\leq j\leq r} g_{j}^{e_{ij}}, i = 0,1,\dots,n-1$.
We can write the result as an $i\times j$ matrix.
\item For each $j$, find an index $i_j$ (denote as $i(j)$) such that $e_{ij}$ is maximum in the $j$-th column.
\item Let $h_j = f_{i(j)}/g_{j}^{e_{i(j)j}}$, take $\beta_j = h_j(\sigma)\alpha^{i(j)}$. Then
$$\beta = \displaystyle\sum_{j=1}^{r} \beta_j$$
is a normal element.
\end{enumerate}

L\"uneburg's algorithm starts with a random element and ends with a normal element. The justification of 
its soundness is as follows:
\begin{Proposition}
L\"uneburg's algorithm always generate a normal element over field $\F_{p^n}$.
\end{Proposition}

\begin{Proof}
First, according to the definition of $T$-Order:
$$f_i = Ord_{\alpha^i} \Leftrightarrow f_i(\sigma)\alpha^i = 0$$

Use Lemma \ref{lem:3}, the minimal/characteristic polynomial for $\sigma$ is $x^n-1$ implies that
any annihilator of $\alpha^i$ ({\it i.e.} $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
of field $\F_{p^n}$, this means $\alpha_i$ are linearly independent with each other ($i=0,1,\dots,n-1$).
Since linear map $f(\sigma)$ guarantees that all elements in the field 
have order $lcm(f_0,\dots,f_{n-1})$, i.e. $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
of elements in set $\{f_0,f_1,\dots,f_{n-1}\}$. This actually corresponds to the assertion in the first step of the algorithm:
$$x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$$

In the second step, after factorization of set $\{f_i\}$, we transform $f_i(\sigma)\alpha^i=0$ to
$$h_j(\sigma)\cdot g_{j}^{e_{i(j)j}}(\sigma) \cdot \alpha^{i(j)} = 0$$

Furthermore we have 
$$g_{j}^{e_{i(j)j}}(\sigma)\beta_j = 0 \Leftrightarrow Ord_{\sigma,\beta_j}(x) = g_{j}^{e_{i(j)j}}(x)$$

Consider the facts: 1) elements in set $\{g_j\}$ are relatively prime; 2) $g_{j}^{e_{i(j)j}}$ is the maximum factor; 
3) $x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$, we deduce that 
$$x^n-1 = \prod_j g_{j}^{e_{i(j)j}}(x) = \prod_jOrd_{\sigma,\beta_j}(x)$$

As a result
$$x^n-1 = Ord_{\sigma,\beta}(x) \implies \beta\text{ is a normal element.}$$
\end{Proof}

L\"uneburg's algorithm is an analytic method, which requires relatively high computational complexity 
mainly because of the factorization in its second step. To overcome the high cost, an inductive method 
is proposed, which is the Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
Before introducing the details of the algorithm, we demonstrate two preliminary lemmas.
\begin{Lemma}
\label{lem:4}
For an arbitrary element $\theta \in \F_{p^n}$ that $Ord_\theta(x) \neq x^n - 1$,
let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.
\end{Lemma}
\begin{Proof}
Assume $\gamma$ is the desired normal element. From the definition of normal element we derive 
$$\exists f(x)\in \F_{p^n}[x],~~f(\sigma)\gamma = \theta$$

Then using the definition of $T$-Order, we derive
$$Ord_{\sigma}\theta = 0 \implies (Ord_{\sigma}f(\sigma))\gamma = 0$$

Since $\gamma$ is the normal element, $Ord_\gamma(x) = x^n-1\implies x^n-1~|~(Ord_\theta(x)f(x))$.
Furthermore,
$$g(x) = \frac{x^n-1}{Ord_\theta(x)} \implies x^n-1\Bigm|\left(\frac{x^n-1}{g(x)}\cdot f(x)\right)\implies g(x)~|~f(x)$$

Let $f(x)=h(x)g(x)$. Then
$$g(\sigma)(h(\sigma)\gamma) = \theta$$

Therefore, $\exists \beta$ {\it i.e. }$g(\sigma)\beta = \theta$. Concretely, $\beta = h(\sigma)\gamma$.
\end{Proof}

\begin{Lemma}
\label{lem:5}
Define $\theta$ and $g(x)$ as in Lemma \ref{lem:4}. Assume there exists a solution 
$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a non-zero element $\eta$ such that
$$g(\sigma)\eta = 0$$
and
$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
\end{Lemma}
\begin{Proof}
Assume $\gamma$ is the desired normal element. Similarly,
$$\exists \eta = Ord_\theta(\sigma)\gamma \neq 0,~~g(\sigma)\eta = 0$$

Follow the setup in Lemma \ref{lem:4}, we have
$$g(\sigma)\beta = \theta,~~\frac{x^n-1}{Ord_\theta(x)}\Bigm|_\sigma\cdot \beta = \theta$$

Subsequently,
$$Ord_\theta(x)\cdot \frac{x^n-1}{Ord_\theta(x)}\Bigm|_\sigma\beta = 0 = Ord_\beta(x)\beta$$
which implies
$$Ord_\theta(x)\bigm|Ord_\beta(x),~~deg(Ord_\theta(x)) \leq deg(Ord_\beta(x))$$
Combining with the assumption in the lemma: $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$, we derive
$$deg(Ord_\beta(x)) = deg(Ord_\theta(x))\implies Ord_\theta(x) =Ord_\beta(x)$$

In the next part we need to prove $g(x)\perp Ord_\theta(x)$ by contradiction. Assume 
$h(x) = gcd(g(x),Ord_\theta(x))\neq 1$, and
$$g(\sigma)\beta = a(\sigma)h(\sigma)\beta = \theta,~~Ord_\theta(\sigma)\theta = a(\sigma)b(\sigma)h^2(\sigma)\beta$$
However, $Ord_\theta(x) =Ord_\beta(x)$ indicates that $Ord_\beta(x)\beta = b(x)h(x)a(x)\beta = 0$. Thus,
$Ord_\theta(x)=b(x)$. This is true if and only if $h(x)=1$. As a result, we assert that
$$g(x) \perp Ord_\theta(x)$$

Consider $g(\sigma)\eta = 0 \implies Ord_\eta(x)\bigm|g(x)$. It further implies that 
$$Ord_\theta(x)\perp Ord_\eta(x) \implies Ord_{\theta+\eta}(x) = Ord_\theta(x)\cdot Ord_\eta(x)$$

Since $\eta \neq 0$, we derive the result
$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
\end{Proof}

Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:\\
\textbf{Lenstra's Algorithm}
\begin{enumerate}[{1)}]
\item Take an arbitrary element $\theta \in \F_{p^n}$, determine $Ord_\theta(x)$.
\item If $Ord_\theta(x) = x^n - 1$ then algorithm terminates and return $\theta$ as a normal element.
\item Otherwise, compute $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.
\item Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to 2);
otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, and go to 2).
\end{enumerate}

Lenstra's algorithm is an approximation algorithm in nature. For each iteration $Ord_\theta(x)$ monotonically increase,
so it will finally reach termination condition $Ord_\theta(x) = x^n-1$.
\subsection{Bases Conversion}
% General \lambda matrix computation

\section{Optimal Normal Basis}
\label{append:ONB}
\subsection{Construction of Optimal Normal Basis}
\subsection{Optimal Normal Basis Multiplier Design}
% Customized Galois Field IEEE1363-2000
% Multi table <=> \lambda matrix
In Section \ref{sec:nbdesign}, we introduce the concept of multiplication table and explained how its 
entries are transformed from the $0$-th $\lambda$-Matrix. Actually we can mathematically prove the relationship 
between them:
\begin{Theorem}
Multiplication table $T$ is actually a conjugate of the $\lambda$-Matrix $M$. 
We can use an equation $M_{i,j}^{(0)} = T_{j-i,-i}$ to demonstrate the relation between their entries.
\end{Theorem}
\begin{Proof}
Recall the definition of $\lambda$-Matrix over field $\F_{2^k}$:
$$C = A\times B = \left(\sum_i a_i\beta^{2^i}\right)\left(\sum_jb_j\beta^{2^j}\right) = \sum_i\sum_ja_ib_j\beta^{2^i}\beta^{2^j}$$
Then there always exist $\lambda_{ij}^{(l)}$, such that 
$$\beta^{2^i}\beta^{2^j} = \sum_l\lambda_{ij}^{(l)}\beta^{2^l}$$
We call it the cross-product term. The $l$-th bit of the product $C$ is
$$c_l = \sum_i\sum_ja_ib_j\lambda_{ij}^{(l)}$$
For the sake of simplification, we let $i=0$, thus 
$$\beta\cdot\beta^{2^j} = \sum_l\lambda_{0j}^{(l)}\beta^{2^l}$$
Since the multiplication table is defined as
\begin{equation}
\beta
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
\end{bmatrix}
= {\bf T}
\begin{bmatrix}
\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
\end{bmatrix}
\end{equation}
The $j$-th row of $T$ can be written as
\begin{align}
\beta\cdot\beta^{2^j} =& \sum_l\lambda_{0j}^{(l)}\beta^{2^l} & \text{(multi-table definition)} \nonumber\\
					=& \sum_l\lambda_{jl}^{(0)}\beta^{2^l} & \text{(cross-product term)} \label{eqn:mteq}
\end{align}
Notice that $\lambda$ corresponds to $\lambda$-Matrix entries. In Equation \ref{eqn:mteq} we assign $0$ to the 
row index $i$, the proof can be extended to all row indices $i<n$ but fix $l$ to 0 because of the 
conjugation generated by right-down cyclic-shift of $\lambda$-Matrix. Therefore we have 
$$M_{i,j}^{(0)} = \lambda_{ij}^{(0)} = \lambda_{j,0}^{(i)} = T_{j-i,0-i}$$
\end{Proof}
